---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading packages
```{r, results='hide',  message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity/activity.csv")
```
## What is mean total number of steps taken per day?

```{r}
activity_by_date <- activity %>%
  group_by(date) %>%
  summarize(total.steps = sum(steps)) 


day_mean <- mean(activity_by_date$total.steps, na.rm=TRUE)
day_median <- median(activity_by_date$total.steps, na.rm=TRUE)
```
The **average number** of steps per day was `r format(day_mean, scientific=FALSE)`, and the **median** was `r format(day_median, scientific=FALSE)`.


```{r histogram, fig.height=5}
p <- ggplot(activity_by_date,aes(total.steps))
p + geom_histogram(fill="darkgreen", colour="green") + labs(title="Histogram of steps per day")
```

## What is the average daily activity pattern?
```{r daily_activity}
average_day <- activity %>%
  group_by(interval) %>%
  summarize(average.steps=mean(steps, na.rm=TRUE))
most_active <- which.max(average_day$average.steps)
most_active_string <- sprintf("%04d",unlist(average_day[most_active, "interval"]))
most_active_string <- paste(substr(most_active_string,1,2), substr(most_active_string,3,4), sep=":")
max_steps <- average_day$average.steps[most_active]
```
```{r interval_plot, fig.height=5}
p <- ggplot(data=average_day, aes(interval, average.steps))
p + geom_point(colour="blue") + geom_line(colour="blue", size=1) +scale_x_continuous(limits=c(0,2355), expand=c(0,0), labels=c("0:00", "6:00", "12:00",  "18:00", "23:00"), breaks=c(0, 600, 1200, 1800, 2300)) +
  labs(title="Average number of steps for a 5 minute interval", x="Time of day", y="Number of steps")
```
  
The most active 5 minute time interval is at `r most_active_string`, with the average numbver of `r round(max_steps,1)` steps.


## Imputing missing values
Missing values for each of the 5 minute intervals are vreploaced with the avreage number of time steps for that time interval calculated aver all days.
```{r input_missing}
activity.na.rm <- activity
index_missing <- which(is.na(activity.na.rm$steps))
for(i in index_missing){
  activity.na.rm[i, "steps"] <- average_day[average_day$interval==activity.na.rm[i,"interval"],"average.steps"]
}
```

```{r}
activity.na.rm_by_date <- activity.na.rm %>%
  group_by(date) %>%
  summarize(total.steps = sum(steps)) 


day_mean.na.rm <- mean(activity.na.rm_by_date$total.steps, na.rm=TRUE)
day_median.na.rm <- median(activity.na.rm_by_date$total.steps, na.rm=TRUE)
```
The **average number** of steps per day was `r format(day_mean.na.rm, scientific=FALSE)`, and the **median** was `r format(day_median.na.rm, scientific=FALSE)`. The average does not change, which makes sense as we input the average for all NA. Since the only records missing were records for entire days, this operation does not change the daily average. The median does change - it is now equal to the mean, which again makes sense as we added quite a few "average" days into our data and the median is likely to be equal to one of them. 


```{r histogram2, fig.height=5}
p <- ggplot(activity.na.rm_by_date,aes(total.steps))
p + geom_histogram(fill="darkgreen", colour="green") + labs(title="Histogram of steps per day, filled NA")
```
    
## Are there differences in activity patterns between weekdays and weekends?

```{r weekday_analysis}
activity.na.rm_by_weekend <- activity.na.rm %>% 
  mutate(day=weekdays(as.Date(date), abbreviate=TRUE)) %>%
  group_by(interval, is.weekend=ifelse(is.element(day,c("Sat", "Sun")), "Weekend", "Weekday")) %>% 
  summarize(average.steps = mean(steps))
```

```{r time_series_panel, fig.height=5}
p <- ggplot(data=activity.na.rm_by_weekend, aes(interval, average.steps))
p + geom_point(colour="blue") + facet_grid(is.weekend~.) + geom_line(colour="blue", size=1) + scale_x_continuous(limits=c(0,2355), expand=c(0,0), labels=c("0:00", "6:00", "12:00",  "18:00", "23:00"), breaks=c(0, 600, 1200, 1800, 2300)) +
  labs(title="Average number of steps for a 5 minute interval on weekends and weekdays.", x="Time of day", y="Number of steps") 

```

